---
title: "R Notebook"
output:
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: inline
---

```{r}
library(tidyverse)
library(broom)
library(ggforce)
library(ggridges)
library(gridExtra)
source("Spatial_indicators_functions_Woillez2009_modified.R")
```


```{r, eval=F}
sst <- readRDS("Data/temperature_data.RDS") %>% 
  inner_join(readRDS("Data/spatial_data.RDS"))

bering <- as.character(c(514,524,508,512,516,509,513,517,521))

sst2 <- sst %>% 
  filter(!is.na(NMFSAREA)) %>% 
  mutate(region=ifelse((NMFSAREA%in%bering) & maxlat>60.1 & maxlat<65.6,"NBS",
                       ifelse((NMFSAREA%in%bering) & maxlat<=60.1,"EBS",
                              ifelse(NMFSAREA%in%c(610),"WGOA",
                                     ifelse(NMFSAREA%in%c(620,630),"CGOA",
                                            ifelse(NMFSAREA%in%c(640,650),"EGOA","OTHER"))))),
         moday=substr(as.character(date),6,10)) %>% 
  dplyr::select(-c(STATEFED,sst.sd,m.depth,sd.depth)) %>% 
  inner_join(readRDS("Data/Stat_Centroid.RDS") %>% 
               mutate(STAT_AREA=as.character(STAT_AREA))) %>% 
  mutate(year=format(date,"%Y"))

saveRDS(sst2,file="Data/temperature_data_spatdat.RDS")
```

Read in the data. 
1) Get rid of areas outside of the Alaska fisherhy management areas.
2) Create a new SST field with positive values only. The coldest value is set at -1.8 so add an offset of 1.8 to this adjusted field. This is helpful for the Inertia calculation which gets thrown off by negatives. Given that the sst value for inertia is only used as a relative weighting factor it doesn't matter what the actual number is for that part. 

```{r}
sst <- readRDS("Data/temperature_data_spatdat.RDS") %>% 
  filter(region!="OTHER") %>% 
  mutate(sstadj=sst.mean+1.8)

head(sst)
```

Let's just explore some general trends in the sst data for each region first. We begin with calculating the daily average of sst across all stat areas within each region. The GAM smoother has a cyclic spline to account for the period nature of the time series.

```{r daily_avg_sst_region,fig.cap="Average daily SST across each region with a GAM smoother"}
meandat <- sst %>% 
  group_by(region,date) %>% 
  summarise(mymean=mean(sst.mean,na.rm=TRUE)) %>% 
  ungroup %>% 
  group_by(region) %>% 
  mutate(index=1:n()) %>% 
  dplyr::select(region,index,mymean) 

meandat %>% 
  ggplot(aes(index,mymean)) + 
  geom_line() + 
  geom_smooth(formula = y ~ s(x, bs = "cc")) + 
  facet_wrap(~region)
```

Stack the smoothing splines for each region.

```{r}
meandat %>% 
  ggplot(aes(index,mymean,color=region)) + 
  geom_smooth(formula = y ~ s(x, bs = "cc")) 
```

Or if we used linear smoothers for each region and plotted the slopes of those smoothers to compare regional warming, we get the following. The x-axis here represents years and shows an average warming of 1 degree C in 10 - 18 or so years, depending on the region. The EBS actually has a greater slope with 9.9 degrees per 100 years versus 8.9 for the NBS.

```{r}
meandat %>% 
  group_by(region) %>% 
  do(tidy(lm(mymean~index,data=.))) %>% 
  filter(term!="(Intercept)") %>% 
  mutate(annualincrease=estimate*365) %>% 
  ggplot() + 
  geom_abline(aes(slope=annualincrease,intercept=0,color=region)) + 
  ylim(0,1) + 
  xlim(0,20)
```

Now look at the center of gravity of the ADFG stat areas within each region, weighted by temperature. This is a relative metric to examine how the location of the thermal mass moves within and across years. The key is that each year has all of the same spatial coordinates. So if the temperature in a particular location is greater, it will increase the weight of the latitude and longitude at that location, thereby shifting the thermal mass in that direction. So if you see the center gravity move in a direction, that is the direction in which the warmer water is moving. 

Note that for these calculations we have to use the temperature field with the 1.8 degree offset.

```{r}
mycgi <- sst %>% 
  group_by(date,region) %>% 
  do((cgi(x=.$x,y=.$y,z=.$sstadj) %>% data.frame)) %>% 
  data.frame

d1 <- mycgi %>% 
  dplyr::select(date,region,xcg,ycg,I) %>% 
  mutate(year=format(date,"%Y"),
         month=as.numeric(format(date,"%m")))
```

Overlay the daily centers of gravity for each area with a gradient for year to see if anything jumps out. 

```{r,fig.cap="Daily center of gravity across days and regions for all years"}
d1 %>% 
  filter(year<2019) %>% 
  ggplot(aes(xcg,ycg,color=as.numeric(year))) + 
  geom_point() + 
  facet_wrap(~region,scales="free") + 
  theme_bw() + 
  theme(legend.position = c(0.8,0.25))
```

Now look at how the center of gravity shifts within a year for each area.

Northern Bering Sea - generally counter-clockwise flow within each year. The biggest standout may be 2018 with it's very tight distribution, likely the result of an absence of sea ice. 

```{r nbs_cgi,fig.cap="Daily center of gravity for the Northern Bering Sea management area"}
d1 %>% 
  filter(region=="NBS" & year<2019) %>% 
  ggplot(aes(xcg,ycg,color=month)) + 
  geom_point() + 
  facet_wrap(~year,ncol=4) + 
  theme_bw() + 
  theme(legend.position = "top")
```

In the above figure we saw that more recently (loss of sea ice?) there was a tighter spatial distribution of the thermal mass throught the year. This should translate into a lower spatial inertia (essentially, a lower standard deviation of lat and long). Let's explore this by looking at the average annual inertia for the NBS relative to the other regions. How much do they all change? Or do they? 

```{r}
d1 %>% 
  distinct() %>% 
  group_by(region,year) %>% 
  summarise(meanI=mean(I,na.rm=TRUE)) %>% 
  ggplot(aes(year,meanI)) + 
  geom_point() + 
  facet_wrap(~region)
```

Damnnnnnnnnn, what is up, NBS? Let's look at this daily now instead of annually to see if things look different. The x-axis in the below figure just represents days since the start of the time series (01-01-2003).

```{r}
d1 %>% 
  distinct() %>% 
  group_by(region) %>% 
  mutate(year=as.numeric(year),
         index=1:n()) %>% 
    ungroup %>% 
  ggplot(aes(index,I,color=year)) + 
  geom_line() + 
  facet_wrap(~region) + 
  theme(legend.position = c(0.8,0.25))
```

Looks like there is a potential relationship between inertia in the NBS and warm vs. cold years. Makes sense - in colder years, there is more ice, which leads to so crazy dynamics during the melting period. 

Let's look at the NBS more specifically. The below figure shows daily inertia values for the NBS within and across years.  
```{r}
d1 %>% 
  distinct() %>% 
  filter(region=="NBS") %>% 
  group_by(year) %>% 
  mutate(index=1:n()) %>% 
    ungroup %>% 
  ggplot(aes(index,year,height=I)) + 
  geom_density_ridges(stat="identity",scale=1) +
  theme_bw() + 
  xlab("Julian day")
```

Might be easier to distill the NBS into monthly figures. So let's calculate a CG for each month
```{r}
mycgimonth <- sst %>% 
  mutate(month=as.numeric(format(date,"%m"))) %>% 
  group_by(year,month,region) %>% 
  do((cgi(x=.$x,y=.$y,z=.$sstadj) %>% data.frame)) %>% 
  data.frame

dmo <- mycgimonth %>% 
  dplyr::select(month,region,xcg,ycg,I,year) 
```

For most regions, a monthly center of gravity and inertia don't seem that interesting except for the nbs.

```{r}
dmo %>% 
  distinct() %>% 
  group_by(year,region) %>% 
  mutate(index=1:n()) %>% 
    ungroup %>% 
  ggplot(aes(index,I)) + 
  geom_line() +
  facet_grid(year~region) +
  theme_bw() + 
  xlab("Julian day")
```


```{r}
dmo %>% 
  distinct() %>% 
  filter(region=="NBS") %>% 
  group_by(year) %>% 
  mutate(index=1:n()) %>% 
    ungroup %>% 
  ggplot(aes(index,year,height=I,fill=month)) + 
  geom_density_ridges_gradient(stat="identity",scale=1) +
  theme_bw() + 
  xlab("Julian day")
```


Eastern Bering Sea

```{r ebs_cgi,fig.cap="Daily center of gravity for the Eastern Bering Sea management area"}
d1 %>% 
  filter(region=="EBS" & year<2019) %>% 
  ggplot(aes(xcg,ycg,color=month)) + 
  geom_point() + 
  facet_wrap(~year,ncol=4) + 
  theme_bw() + 
  theme(legend.position = "top")
```

So 2012 seems to really stick out. What does the annual cg and inertia look like for the EBS - does 2012 stick out there? The below figure was meant to explore the difference between 2012 and other years but I think that because 2012's variability in cg is relatively symmetrical, perhaps there is a canceling effect such that it doesn't really stick out based on cg and inertia. Perhaps GIC will be better?

```{r}
mycgiyear <- sst %>% 
  group_by(year,region) %>% 
  do((cgi(x=.$x,y=.$y,z=.$sstadj) %>% data.frame)) %>% 
  data.frame

dyr <- mycgiyear %>% 
  dplyr::select(region,xcg,ycg,I,year) %>% 
  distinct()

dyr %>% 
  ggplot(aes(as.numeric(year),I)) + 
  geom_line() +
  geom_point() +
  facet_wrap(~region,scales="free") +
  theme_bw() + 
  xlab("Julian day")
```

Western GOA seems remarkably unremarkable after the Bering Sea plots. Not a lot of heterogeneity. Mostly oriented from lower left (SW) towards upper right (NE) of the coordinate space

```{r wgoa_cgi,fig.cap="Daily center of gravity for the western GOA management area"}
d1 %>% 
  filter(region=="WGOA" & year<2019) %>% 
  ggplot(aes(xcg,ycg,color=month)) + 
  geom_point() + 
  facet_wrap(~year,ncol=4) + 
  theme_bw() + 
  theme(legend.position = "top")
```

Central GOA is also pretty unremarkable. 

```{r cgoa_cgi,fig.cap="Daily center of gravity for the central GOA management area"}
d1 %>% 
  filter(region=="CGOA" & year<2019) %>% 
  ggplot(aes(xcg,ycg,color=month)) + 
  geom_point() + 
  facet_wrap(~year,ncol=4) + 
  theme_bw() + 
  theme(legend.position = "top")
```

Eastern GOA is most interesting probably in that the orientation is opposite that of the other GOA regions. How do the prevailing current directions differ in the egoa compared to the wester and centeral GOA?

```{r egoa_cgi,fig.cap="Daily center of gravity for the Eastern GOA management area"}
d1 %>% 
  filter(region=="EGOA" & year<2019) %>% 
  ggplot(aes(xcg,ycg,color=month)) + 
  geom_point() + 
  facet_wrap(~year,ncol=4) + 
  theme_bw() + 
  theme(legend.position = "top")
```

OKay, so it's difficult to really compare how an individual year compares to the other years with the above metric. Let's plot annual cg's and inertia using ellipses. 

```{r}
mycgiyear <- sst %>% 
  group_by(year,region) %>% 
  do((cgi(x=.$x,y=.$y,z=.$sstadj) %>% data.frame)) %>% 
  data.frame

#mycgifun <- function(mycgi){
#  return(bind_cols(mycgi %>% 
#            dplyr::select(year,xaxe1,xaxe2) %>% 
#            gather(xaxis,xval,-c(year)) %>% 
#            dplyr::select(-xaxis),
#          mycgi %>% dplyr::select(year,yaxe1,yaxe2) %>% 
#            gather(yaxis,yval,-year) %>% 
#            dplyr::select(-yaxis,-year)))
#}

mycgifun <- function(mycgi){
  return(bind_cols(mycgi %>% 
                     dplyr::select(year,xaxe1,xaxe2,region) %>% 
                     gather(xaxis,xval,-c(year,region)) %>% 
                     dplyr::select(-xaxis),
                   mycgi %>% dplyr::select(year,yaxe1,yaxe2,region) %>% 
                     gather(yaxis,yval,-c(year,region)) %>% 
                     dplyr::select(-yaxis,-year)))
}

mycgifun(mycgiyear) %>% 
  filter(!is.nan(xval) & year<2019) %>% 
  ggplot(aes(xval,yval,fill=factor(year))) +
  geom_mark_ellipse(expand = unit(0, "mm"),alpha=0.1) +
  facet_wrap(~region,scales="free") + 
  theme_bw()
```


Plot it slightly differently

```{r}
plotellipsefun <- function(myregion){
  print(mycgifun(mycgiyear) %>%
          filter(!is.nan(xval) & year<2019 & region==myregion) %>% 
          ggplot(aes(xval,yval,fill=factor(year))) +
          geom_mark_ellipse(expand = unit(0, "mm"),alpha=0.1) +
          theme_bw() + 
          theme(legend.position = "none",
                axis.title = element_blank()))
}

grid.arrange(plotellipsefun("WGOA"),
             plotellipsefun("CGOA"),
             plotellipsefun("EGOA"),
             plotellipsefun("NBS"),
             plotellipsefun("EBS"))
```

Okay, let's explore global indices of collocation now to quantify the similarity of one year to another. 

```{r, eval=FALSE}
regionalgicfun <- function(myregion){
  myyear <- 2003:2018
  sstout <- list()
    for(i in 1:length(myyear)){
      sstout[[i]] <- sst %>% 
        filter(region==myregion) %>% 
        group_by(moday) %>% 
        summarise(gic=gic(x1=x[year==myyear[i]],
                          y1=y[year==myyear[i]],
                          z1=sstadj[year==myyear[i]],
                          x2=x[year!=myyear[i]],
                          y2=y[year!=myyear[i]],
                          z2=sstadj[year!=myyear[i]]),
                  year=myyear[i])
    }
  return(sstout)
}

gic1 <- regionalgicfun("NBS") %>% bind_rows() %>% mutate(region="NBS")
gic2 <- regionalgicfun("EBS") %>% bind_rows() %>% mutate(region="EBS")
gic3 <- regionalgicfun("WGOA") %>% bind_rows() %>% mutate(region="WGOA")
gic4 <- regionalgicfun("CGOA") %>% bind_rows() %>% mutate(region="CGOA")
gic5 <- regionalgicfun("EGOA") %>% bind_rows() %>% mutate(region="EGOA")

saveRDS(bind_rows(gic1,gic2,gic3,gic4,gic5),file="Regional_GIC_Annual_Comparisons.RDS")
```


```{r}
data <- readRDS("Regional_GIC_Annual_Comparisons.RDS")

data %>%
  group_by(year,region) %>% 
  mutate(index=1:n()) %>% 
  ungroup %>% 
  ggplot(aes(index,gic)) + 
  geom_line() + 
  facet_grid(region~year,scales="free")

```

